Bayesian multivariate skew-normal finite mixture model for analysis of infant development trajectories

### 2019-06-24 manuscript/manuscript.tex
section 3

\subsection{Generic Multivariate Skew Normal Mixture}

Let $\mathbf{y}_{i}$ be the $J \times 1$ observation vector for subject $i$ such that $y_{ij}$ is the observation for subject $i$ at timepoint $j$. We assume for now that $\mathbf{y}_{i}$ is fully observed. Later, we relax this assumption by allowing for missingness in the components of $\mathbf{y}_{i}$. For the analysis of the Nurture data, we propose the following MSN mixture
\begin{equation}
f(\mathbf{y}_i) = \sum_{k = 1}^{K} \pi_{ik} f(\mathbf{y}_i|\boldsymbol\theta_k),
\end{equation}
where $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$ ($k = 1,...,K$). For now we assume that $K$ is fixed, but we discuss model selection strategies for choosing the optimal value of $K$ in Section 4. The probability that subject $i$ ($i = 1,...,n$) belongs to cluster $k$ is denoted $\pi_{ik}$. Note that $\pi_{ik}$ is indexed by $i$, as we allow the probability of class membership to vary across subjects. However, we assume a that class membership is fixed throughout the study period, since our focus is to cluster individuals based on their overall developmental patterns over the course of the study. In Section 6, we discuss extensions to allow for class membership to vary over time.

In each cluster, we assume that the $J\times 1$ outcome vector for subject $i$ follows an MSN distribution. Let $z_i \in \{1,...,K\}$ denote a latent cluster membership of subject $i$. Conditional on $z_i = k$, we model $\mathbf{y}_{i}$ as 
\begin{equation}
\mathbf{y}_{i}|z_i=k \sim MSN_J(\boldsymbol\zeta_k,\boldsymbol\alpha_k,\boldsymbol\Omega_k),
\end{equation}
where $\boldsymbol\zeta_k$ is a $J \times 1$ vector of cluster-specific location parameters, $\boldsymbol\alpha_k$ is a $J \times 1$ vector of cluster-specific skewness parameters, and $\boldsymbol\Omega_k$ is a $J \times J$ cluster-specific scale matrix that accounts for associations between the $J$ responses. The vector $\boldsymbol\alpha_k$ has components $\alpha_{jk}$, $j = 1,...,J$, that control the skewness of outcome $j$ in cluster $k$. When $\boldsymbol\alpha_k = \mathbf{0}$, the MSN distribution reduces to the multivariate normal distribution $N_J(\boldsymbol\zeta_k,\boldsymbol\Sigma_k)$, where $\boldsymbol\Sigma_J$ is the $J \times J$ covariance matrix capturing dependence among the $J$ outcomes.

\subsection{Multivariate Skew Normal Stochastic Representation}

We model the effect of covariates on longitudinal motor development outcomes through the use of a MSN regression model. We adopt a standard stochastic representation of a MSN random variable by conditioning on a subject-specific latent truncated normal random effect, denoted $t_i$ (Fr\"{u}hwirth-Schnatter and Pyne 2010). Conditional on membership in cluster $k$, the stochastic representation of the MSN takes the form
\begin{equation}
\mathbf{y}_{ik} = \mathbf{x}_{ik}^T\boldsymbol\beta_k  + t_{i}\boldsymbol\psi_k + \boldsymbol\epsilon_{ik}
\end{equation}

where $\mathbf{x}_i$ is the $P \times 1$ vector of covariate values for subject $i$, $\boldsymbol\beta$ is the $P \times J$ vector of fixed effects coefficients , $t_i \stackrel{iid}{\sim} N_{[0,\infty)}(0,1)$ is a conditional truncated normal random effect, $\boldsymbol\psi$ is the $J \times 1$ vector containing cluster-specific skewness parameters, and $\boldsymbol\epsilon_i \sim N_J(0,\boldsymbol\Sigma)$ is the correlated error term. In can be shown that under these conditions $\mathbf{y}_i$ is multivariate skew normal with location. For conciseness, we present the MSN distribution here outside of the finite mixture context. When using this stochastic representation in our finite mixture model, the parameters $\boldsymbol\beta$, $\boldsymbol\psi$, and $\boldsymbol\Sigma$ are replaced with cluster-specific counterparts $\boldsymbol\beta_k$, $\boldsymbol\psi_k$, and $\boldsymbol\Sigma_k$.

\subsection{Representation as Matrix Skew Normal}

We note that the multivariate skew normal regression problem can be expressed more succinctly through the use of Matrix Skew Normal distributions. In particular, we define $\mathbf{Y}_{n_k \times J}$ as the response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n \mathds{1}_{(z_i = k)}$ is the total number of observations belonging to cluster $k$ ($k = 1,...,K$). 
$$\mathbf{Y}_{n_k \times J} \sim MatSN_{n_k \times J}(\mathbf{M},\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Sigma_k)$$
$$\text{vec}(\mathbf{Y}) = \mathbf{y}_{n_kJ \times 1} = (\mathbf{y}^T_1,...,\mathbf{y}^T_{n_k})$$
$$\text{vec}(\mathbf{M}) = \boldsymbol\mu_{n_k J \times 1} = (\boldsymbol\mu_1^T,...,\boldsymbol\mu_{n_k}^T)$$

where $\boldsymbol\mu_i = \mathbf{x}_i \boldsymbol\beta_k$ and $\boldsymbol\alpha_k = (\alpha_{k1},...,\alpha_{kJ})$. This implies that the $i^{th}$ row of $\mathbf{Y}$ is $\mathbf{y}_i \sim MSN_J(\boldsymbol\mu_i, \boldsymbol\alpha_k, \boldsymbol\Sigma_k) = {MSN}_J(\mathbf{x}_i \boldsymbol\beta_k, \boldsymbol\alpha_k, \boldsymbol\Sigma_k)$. Using the stochastic representation presented in Section 3.2, the conditional distribution of $\mathbf{y}_i | z_i, t_i$ is $N_J (\mathbf{x}_i \boldsymbol\beta_k + \boldsymbol\psi_k t_i,\boldsymbol\Sigma_k)$. This implies that all responses for cluster $k$ conditioned on $\mathbf{T}_{n_k \times 1}$ is distributed as 
$$\mathbf{Y}_{n_k \times J} | \mathbf{T}_k = \mathbf{t}_k \sim MatNorm_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k)$$
$$\text{vec}(\mathbf{M}_k)_{n_kJ \times 1} = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$$\

where $\mathbf{t}_k = (t_1,...,t_{n_k})$.

\subsection{Multinomial Regression on Cluster Probabilities}

A primary concern of our model is with identification of latent infant development classs. We accomplish this via multinomial logit regression model on cluster membership, which utilizes P\'{o}lya-Gamma data-augmentation, as described in Section 3.6 to allow for updating of all parameters using Gibbs sampling. The multinomial logit model is as follows for $k = 1,...,K$.
$$P(z_i = k|\mathbf{w}_i) = \pi_{ik} = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k}}{\sum_{k' = 1}^K e^{\mathbf{w}_i^T \boldsymbol\delta_{k'}}}$$

where $\mathbf{w}_i$ is the vector of cluster probability covariates for subject $i$, $\boldsymbol\delta_k$ contains the multinomial regression parameters for cluster $k$, and $K$ is the number of putative classs specified \textit{a priori}. For identifiability purposes, we fix the reference category $k = K$ and set $\boldsymbol\delta_K = \mathbf{0}$. Under this model, $z_i|\mathbf{w}_i \sim Multinom(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$. During MCMC estimation, the cluster labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as cluster assignments.

\subsection{Conditional MVN Imputation}

We allow for missingness of outcomes in the MSN mixture model by imputing missing values from their conditional multivariate normal distributions. We note that
$$\mathbf{y}_i|z_i,\mathbf{x}_i,t_i,\boldsymbol\beta_{k},\boldsymbol\psi_{k},\boldsymbol\Sigma_{k} \sim N_J(\boldsymbol\beta_{k}^T\mathbf{x}_i  + t_i \boldsymbol\psi_{k}, \boldsymbol\Sigma_{k})$$

where $z_i = k$. This allows us to appeal to standard conditional forms of the multivariate normal distribution to specify the distribution of missing observations. For subject $i$, we use $q_i$ to denote the number of the total $J$ possible repeated measurements are missing. We partition the full $J \times 1$ outcome vector $\mathbf{y}_i$ into the $(J - q_i) \times 1$ observed data vector $\mathbf{y}_i^{obs}$ and the $q_i \times 1$ missing data vector $\mathbf{y}_i^{miss}$. $Y_i = [Y^{miss}_{i_{q \times 1}} | Y^{obs}_{i_{J - q \times 1}}]^T$. We have
$$\mathbf{y}_i^{miss}|\mathbf{y}_i^{obs},\mathbf{x}_i,t_i,z_i,\boldsymbol\beta_{k},\boldsymbol\psi_{k},\boldsymbol\Sigma_{k} \sim N(\boldsymbol\mu^{miss},\boldsymbol\Sigma^{miss})$$

where $\boldsymbol\mu^{miss}$ and $\boldsymbol\Sigma^{miss}$ take standard forms detailed in the Appendix.

An attractive feature of this imputation approach is that each missing outcome is imputed ``online", i.e. once per MCMC iteration. This provides more opportunities to explore the parameter space than multiple imputation and avoids multiplicative run-time scaling in $m$, the number of imputations. We demonstrate this feature using simulations detailed in Section 4.

@@

\subsubsection{P\'olya--Gamma Data Augmentation}

Polson \textit{et al.} (2013) introduced an efficient data augmentation approach to fitting several GLMs including the multinomial logit regression model specified by Holmes and Held (2006), which specifies the full conditional distributions of the multinomial regression parameters as a function of a Bernoulli likelihood. 
$$p(\boldsymbol\delta_k|\mathbf{z},\boldsymbol\delta_{k' \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ik}^{U_{ik}}(1-\pi_{ik})^{1-U_{ik}}$$

where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $U_{ik} = \mathds{1}_{z_i = k}$ is an indicator that subject $i$ belongs to cluster $k$, and $\pi_{ik}$ is defined as in Section 3.4. We can rewrite $\pi_{ik}$ as follows
$$\pi_{ik} = P(U_{ik} = 1) = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}}{1 + e^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}} = \frac{e^{\eta_{ik}}}{1 + e^{\eta_{ik}}}$$

@@

\subsubsection{Prior Choice}

@@

\caption{\label{tab:unnamed-chunk-4}Model results for simulated data with n = 1500, k = 4, p = 1, h = 3, v = 1. 5000 iterations were run with a burn in of 1000. Missingness mechanism was MAR and P(miss) = 0}
\centering
\fontsize{8}{10}\selectfont
\begin{tabular}{llllllll}
\toprule
\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Cluster 1} & \multicolumn{2}{c}{Cluster 2} & \multicolumn{2}{c}{Cluster 3} \\
\cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6} \cmidrule(l{3pt}r{3pt}){7-8}
Model Component & Parameter & True & Est. (95\% CrI) & True & Est. (95\% CrI) & True & Est. (95\% CrI)\\
\midrule
\addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{ }}\\
\hspace{1em}MVSN & $\beta_{0}$ & -1.35 & -1.59 (-2.3, -0.94) & -0.18 & -0.24 (-0.53, 0.38) & 0.72 & 0.79 (0.52, 1.02)\\
\hspace{1em}Regression & $\beta_{1}$ & -1.19 & -1.35 (-1.99, -0.84) & -0.09 & -0.33 (-0.62, 0.78) & 1.64 & 1.65 (1.33, 1.91)\\
\hspace{1em} & $\beta_{2}$ & -1.65 & -1.81 (-2.66, -1.41) & -0.47 & -0.62 (-0.89, 0.3) & 1.44 & 1.32 (0.97, 1.6)\\
\hspace{1em} & $\beta_{3}$ & -1.75 & -1.89 (-2.52, -1.37) & -0.22 & -0.32 (-0.63, 0.55) & 2.28 & 2.26 (1.96, 2.5)\\
\addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{ }}\\
\hspace{1em} & $\sigma_{11}$ & 1 & 1.02 (0.78, 1.19) & 1 & 0.98 (0.77, 1.23) & 1 & 1.06 (0.85, 1.29)\\
\hspace{1em} & $\sigma_{12}$ & -0.32 & -0.19 (-0.33, -0.02) & 0.16 & 0.14 (-0.01, 0.4) & 0.72 & 0.82 (0.62, 1.05)\\
\hspace{1em} & $\sigma_{13}$ & -0.65 & -0.55 (-0.68, -0.35) & 0.72 & 0.7 (0.51, 0.94) & 0.14 & 0.27 (0.1, 0.48)\\
\hspace{1em} & $\sigma_{14}$ & -0.44 & -0.33 (-0.46, -0.13) & 0.5 & 0.48 (0.31, 0.72) & -0.01 & -0.02 (-0.16, 0.16)\\
\hspace{1em} & $\sigma_{22}$ & 1 & 0.92 (0.72, 1.06) & 1 & 0.94 (0.72, 1.22) & 1 & 1.11 (0.87, 1.38)\\
\hspace{1em} & $\sigma_{23}$ & 0.56 & 0.49 (0.33, 0.6) & 0.53 & 0.46 (0.29, 0.73) & -0.1 & 0.08 (-0.1, 0.28)\\
\hspace{1em} & $\sigma_{24}$ & 0.98 & 0.9 (0.7, 1.04) & 0.24 & 0.14 (-0.03, 0.41) & 0.19 & 0.17 (0.01, 0.37)\\
\hspace{1em} & $\sigma_{33}$ & 1 & 0.9 (0.66, 1.04) & 1 & 0.92 (0.72, 1.19) & 1 & 1.26 (1.03, 1.52)\\
\hspace{1em} & $\sigma_{34}$ & 0.56 & 0.51 (0.35, 0.62) & 0.86 & 0.79 (0.58, 1.05) & -0.65 & -0.63 (-0.78, -0.45)\\
\hspace{1em} & $\sigma_{44}$ & 1 & 0.93 (0.72, 1.07) & 1 & 0.93 (0.7, 1.2) & 1 & 1.11 (0.87, 1.36)\\
\addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{ }}\\
\hspace{1em} & $\psi_{1}$ & -0.33 & -0.02 (-0.84, 0.87) & 0.67 & 0.69 (-0.09, 1.01) & -1 & -0.98 (-1.25, -0.67)\\
\hspace{1em} & $\psi_{2}$ & -0.33 & -0.16 (-0.8, 0.62) & 0.67 & 0.81 (-0.6, 1.14) & -1 & -0.98 (-1.28, -0.6)\\
\hspace{1em} & $\psi_{3}$ & -0.33 & -0.15 (-0.65, 0.89) & 0.67 & 0.72 (-0.35, 1.02) & -1 & -0.8 (-1.15, -0.36)\\
\hspace{1em} & $\psi_{4}$ & -0.33 & -0.18 (-0.84, 0.59) & 0.67 & 0.71 (-0.32, 1.06) & -1 & -1 (-1.32, -0.62)\\
\addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{ }}\\
\hspace{1em}Multinom. & $\delta_{11}$ & -0.84 & -0.78 (-0.96, -0.59) & -0.84 & -0.78 (-0.96, -0.59) & -0.84 & -0.78 (-0.96, -0.59)\\
\hspace{1em} & $\delta_{12}$ & -0.24 & -0.26 (-0.42, -0.1) & -0.24 & -0.26 (-0.42, -0.1) & -0.24 & -0.26 (-0.42, -0.1)\\
\addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{ }}\\
\hspace{1em}Clustering & $\pi_l$ & 0.39 & 0.39 (0.38, 0.4) & 0.26 & 0.26 (0.25, 0.27) & 0.34 & 0.35 (0.33, 0.36)\\
\bottomrule
\end{tabular}
\end{table}

>>>>

\subsection{Multivariate Skew Normal Mixture Model}

A primary goal of the Nurture study is to identify clusters of infants characterized by distinct motor development trajectories. To address this aim, we propose a flexible finite mixture model that accommodates relevant features of the data, such as skewness and dependence among the responses. To this end, let $\mathbf{y}_{i}=(y_{i1},\ldots,y_{iJ})^T$ be a $J \times 1$ vector of responses (i.e., Baley scores) for subject $i$ $(i=1,\ldots,n)$. For the analysis of the Nurture data, we propose a finite mixture model of the form
\begin{equation}
f(\mathbf{y}_i) = \sum_{k = 1}^{K} \pi_{ik} f(\mathbf{y}_i|\boldsymbol\theta_k),
\end{equation}
where $\boldsymbol\theta_k$ is the set of parameters specific to cluster $k$ ($k = 1,...,K$) and $\pi_{ik}$ is a subject-specific mixing weight representing the probability that subject $i$ belongs to cluster $k$. For now we assume that $K$ is fixed; in Section 4, we discuss model selection strategies for choosing the optimal value of $K$. We also assume that class membership is fixed throughout the study period, since our focus is to cluster individuals based on their overall developmental patterns over the course of the study. In Section 6, we discuss extensions to allow for class membership to vary over time. \textbf{[We could omit these last two sentence -- are they really needed? Not sure. Maybe keep for now and think about it.]}

To facilitate posterior inference, we introduce a latent cluster indicator variable $z_i$ taking the value $k \in \{1,...,K\}$ with probability $\pi_{ik}$. Conditional on $z_i = k$, we assume $\mathbf{y}_{i}$ is distributed as
\begin{equation}
\mathbf{y}_{i}|(z_i=k) \sim MSN_J(\boldsymbol\zeta_{ki},\boldsymbol\alpha_k,\boldsymbol\Omega_k), \label{eq:msndens}
\end{equation}
where $MSN_J(\cdot)$ denotes the $J$-dimensional multivariate skew normal density, $\boldsymbol\zeta_{ki}$ is a $J \times 1$ vector of subject- and cluster-specific location parameters, $\boldsymbol\alpha_k$ is a $J \times 1$ vector of cluster-specific skewness parameters, and $\boldsymbol\Omega_k$ is a $J \times J$ cluster-specific scale matrix that captures dependence among the $J$ responses. The vector $\boldsymbol\alpha_k$ has components $\alpha_{kj}$, $j = 1,...,J$ \textbf{[let's use $kij$ as our index hierarchy. Please review throughout]}, that control the skewness of outcome $j$ in cluster $k$. When $\boldsymbol\alpha_k = \mathbf{0}$, the MSN distribution reduces to the multivariate normal distribution $N_J(\boldsymbol\zeta_k,\boldsymbol\Omega_k)$, where $\boldsymbol\Omega_k$ is a $J \times J$ covariance matrix.

We can extend model (\ref{eq:msndens}) to the regression setting by modeling $\boldsymbol\zeta_{ki}$ as a function of covariates. Here we adopt a convenient stochastic representation of the MSN density (Azzalini and Dalla Valle, 1996):
\begin{equation}
\mathbf{y}_i|(z_i=k,t_i) = \mathbf{X}_i \boldsymbol\beta_k + t_i \boldsymbol\psi_k + \boldsymbol\epsilon_{ki}, \label{eq:msnreg}
\end{equation}
where 
$\mathbf{X}_i$ is a $J \times Jp$ design matrix that includes potential time-varying covariates (e.g., indicators denoting quarterly visits); $\boldsymbol{\beta}_k=(\beta_{k11},\ldots,\beta_{k1p},\ldots,\beta_{kJ1},\ldots,\beta_{kJp})^T$ is a $Jp\times 1$ vector of cluster- and outcome-specific regression coefficients; $t_i\sim N_{[0,\infty)}(0,1)$ is a subject-specific standard normal random variable truncated below by zero; $\boldsymbol\psi_k=(\psi_{k1},\ldots,\psi_{kJ})^T$ is a $J \times 1$ vector of cluster-specific skewness parameters; and $\boldsymbol{\epsilon}_{ki} \sim N_J(\boldsymbol0,\boldsymbol\Sigma_k)$ is a $J\times 1$ vector of error terms. Thus, conditional on $t_i$ and $z_i=k$, $\boldsymbol{y}_i$ is distributed as $N_J(\mathbf{X}_i \boldsymbol\beta_k + t_i \boldsymbol\psi_k, \boldsymbol{\Sigma}_k)$. Marginally (after integrating over $t_i$), $\mathbf{y}_i$ is distributed $MSN_J(\boldsymbol\zeta_{ik}, \boldsymbol\alpha_k, \boldsymbol\Omega_k)$, where through back-transformation \textbf{[Carter: Please review these expressions carefully and make sure they conform to those in FS and Pyne]}
\begin{eqnarray*}
\boldsymbol\zeta_{ik} &=& \mathbf{X}_i\boldsymbol\beta_k\\
\boldsymbol\alpha_k &=& \frac{1}{\sqrt{1 - \boldsymbol\psi_k^T \boldsymbol\Omega^{-1}_k\boldsymbol\psi_k}} \boldsymbol\Omega^{-1}_k\boldsymbol\psi_k~~\text{and}\\
\boldsymbol\Omega_k &=& \boldsymbol\Sigma_k  + \boldsymbol\psi_k \boldsymbol\psi_k^T.
\end{eqnarray*}
Additional details can be found in Fr\:{u}wirth-Schnatter and Pyne (2010).

Of note, the MSN density can be expressed more compactly in terms of the matrix skew normal (MatSN) density (Chen and Gupta 2005). As we will see in Section 3.6, the matrix representation of the MSN distribution admits convenient conjugate prior distributions for the regression parameters and scale matrices, which in turn leads to efficient Gibbs sampling for posterior inference. Let $\mathbf{Y}_k$ be an ${n_k \times J}$ response matrix with rows $\mathbf{y}_i^T$, $(i = 1,...,n_k)$, where $n_k = \sum_{i = 1}^n {1}_{(z_i = k)}$ is the number of observations in cluster $k$. From equation (\ref{eq:msnreg}), it follows that $\mathbf{Y}_k$ is distributed as
\begin{eqnarray*}
\mathbf{Y}_{k} &\sim& MatSN_{n_k \times J}(\mathbf{M}_k,\boldsymbol\alpha_k,\mathbf{I}_{n_k},\boldsymbol\Omega_k)\\
%\text{vec}(\mathbf{Y}_k) =  (\mathbf{y}^T_1,...,\mathbf{y}^T_{n_k})^T\\
\text{vec}(\mathbf{M}_k) &=& (\boldsymbol\zeta_{k1}^T,...,\boldsymbol\zeta_{kn_k}^T)^T,
\end{eqnarray*}
where $\boldsymbol\zeta_{ki} = \mathbf{X}_i \boldsymbol\beta_k$ as in equation (\ref{eq:msnreg}), $\boldsymbol\alpha_k = (\alpha_{k1},...,\alpha_{kJ})^T$, $\mathbf{I}_{n_k}$ is the $n_k\times n_k$ identity matrix, and $\boldsymbol\Omega_k$ is the $J\times J$ scale matrix defined above in equation (\ref{eq:msndens}). From equation (\ref{eq:msnreg}), it follows that $\mathbf{Y}_{k}$, conditional on the $n_k\times 1$ vector of random effects $\mathbf{t}_{k}$, is jointly distributed in matrix form as
\[
\mathbf{Y}_k | \mathbf{t}_k \sim MatNorm_{n_k \times J}(\mathbf{M}_k, \mathbf{I}_{n_k}, \boldsymbol\Sigma_k),
\]
where $MatNorm_{n_k \times J}(\cdot)$ denotes a $n_k\times J$ matrix normal density, $\text{vec}(\mathbf{M}_k) = \mathbf{X}_{k} \boldsymbol\beta_k + \mathbf{t}_k \otimes \boldsymbol\psi_k$ is an $n_kJ\times 1$ mean vector, $\mathbf{X}_k$ is an $n_kJ \times Jp$ design matrix, $\boldsymbol\beta_k$ is the $(Jp)\times 1$ vector of regression coefficients defined in equation (\ref{eq:msnreg}), and $\boldsymbol\Sigma_k$ is the $J\times J$ conditional covariance of $\boldsymbol\epsilon_{ik}$ given in equation (\ref{eq:msnreg}).

\subsection{Multinomial Regression for the Cluster Indicators}
To accommodate heterogeneity in the cluster-membership probabilities, we model $\pi_{ik}$ as a function of coovariates using a multinomial logit model 
\begin{eqnarray}
\pi_{ik} = \Pr(z_i = k|\mathbf{w}_i) = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k}}{\sum_{h = 1}^K e^{\mathbf{w}_i^T \boldsymbol\delta_{h}}},~ k=1,\ldots,K,
\end{eqnarray}
where $\mathbf{w}_i$ is an $r\times 1$ vector of subject-level covariates, $\boldsymbol\delta_k$ is a $r\times 1$ vector of regression parameters associated with membership in cluster $k$. For identifiability purposes, we fix the reference category $k = K$ and set $\boldsymbol\delta_K = \mathbf{0}$. Under this model, $z_i|\mathbf{w}_i \sim Multinom(1,\boldsymbol\pi_i)$, where $\boldsymbol\pi_i = (\pi_{i1},...,\pi_{1K})$. During MCMC estimation, the cluster labels $z_i$ are updated from their multinomial full conditional distribution and used in the remaining MCMC steps as cluster assignments.

\subsection{Conditional MSN Imputation}
To accommodate missing at random (MAR) responses, we propose a convenient imputation algorithm that can be implemented ``online'' as part of the Gibbs sampler. In Section 6, we discuss extension to allow for non-ignorable missingness. Suppose $\mathbf{y}_i$ has $q_i\in (1,\ldots,J)$ observed values, denoted $\mathbf{y}^{obs}_i$, and $J-q_i$ intermittent missing values, denoted $\mathbf{y}^{miss}_i$. We can use of the stochastic representation given in equation (\ref{eq:msnreg}) to impute $\mathbf{y}^{miss}_i$ from its conditional multivariate normal distribution given $(z_i,t_i,Y^{obs}_i)$:
\begin{eqnarray}
\mathbf{y}^{miss}_i|(z_i=k,t_i,\mathbf{y}^{obs}_i)&\sim& N_{J-q}(\boldsymbol\mu^{miss}_{ik},\boldsymbol\Sigma^{miss}_k),~\text{where}\nonumber\\
\boldsymbol\mu^{miss}_{ik}&=&\\
\boldsymbol\Sigma^{miss}_k&=& \nonumber
\end{eqnarray}
\textbf{Carter -- work on the above -- you will need to define notation and refer readers back to equation (3) as needed. We can discuss next week if needed}
These results follow from conventional multivariate normal theory. An attractive feature of this imputation algorithm is that it provides more opportunities to explore the parameter space than multiple imputation \textbf{[based on summary stats right?]} and avoids multiplicative run-time scaling in $m$, the number of imputations \textbf{Give refs}. In Section 4, we conduct simulation studies to demonstrate that imputing the missing MSN responses improves inferences over complete case analysis.

@@

\subsubsection{Prior Specification}
We adopt a fully Bayesian inferential approach and assign prior distributions to all model parameters. Conveniently, all parameters admit conditionally conjugate priors, which greatly improves posterior computation via a data-augmented Gibbs sampler. For \textbf{[Give priors for each parameter. Be clear about the conditionally joint prior for $\boldsymbol\beta_k$ and $\boldsymbol\Sigma_k$. Where appropriate, explain advantages]}

\subsubsection{Posterior Inference} 
The above prior specification induces closed-form full conditionals that can be efficiently updated as part of a Gibbs sampler outlined below. Additional details, including derivations can be found in the Web Appendix. \textbf{[Think about the best way to organize this section. Maybe see my Bayesian Analysis paper for guidance? We can discuss next week.]}

\paragraph{P\'olya--Gamma Data Augmentation for $z_i$.} The sampler begins by updating the latent cluster indicators $z_i$ $(i=1,\ldots,n)$ from its multinomial logit full conditional. To facilitate sampling, we adopt an efficient data-augmentation approach introduced by Polson \textit{et al.} (2013), which expresses the inverse-logit function as a mixture P\'olya--Gamma densities. \textbf{[See my Bayesian Analysis paper for guidance on this part]}.

\textbf{[I stopped here]}

$$p(\boldsymbol\delta_k|\mathbf{z},\boldsymbol\delta_{k' \ne k}) \propto p(\boldsymbol\delta_k) \prod_{i = 1}^{n} \pi_{ik}^{U_{ik}}(1-\pi_{ik})^{1-U_{ik}}$$

where $p(\boldsymbol\delta_k)$ denotes the prior distribution of $\boldsymbol\delta_k$, $U_{ik} = {1}_{z_i = k}$ is an indicator that subject $i$ belongs to cluster $k$, and $\pi_{ik}$ is defined as in Section 3.4. We can rewrite $\pi_{ik}$ as follows
$$\pi_{ik} = P(U_{ik} = 1) = \frac{e^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}}{1 + e^{\mathbf{w}_i^T \boldsymbol\delta_k - {c}_{ik}}} = \frac{e^{\eta_{ik}}}{1 + e^{\eta_{ik}}}$$

@@

\end{itemize}
\end{document}

@@

\caption{\label{tab:unnamed-chunk-4}Model results for simulated data with n = 1500, k = 4, p = 1, h = 3, v = 1. 5000 iterations were run with a burn in of 1000. Missingness mechanism was MAR and P(miss) = 0}
\centering
\fontsize{8}{10}\selectfont
\begin{tabular}{llllllll}
\toprule
\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Class 1} & \multicolumn{2}{c}{Class 2} & \multicolumn{2}{c}{Class 3} \\
\cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6} \cmidrule(l{3pt}r{3pt}){7-8}
Model Component & Parameter & True & Est. (95\% CrI) & True & Est. (95\% CrI) & True & Est. (95\% CrI)\\
\midrule
\addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{ }}\\
\hspace{1em}MVSN & $\beta_{0}$ & -3.21 & -3.34 (-3.8, -2.99) & -0.32 & -0.33 (-0.48, -0.14) & 3.35 & 3.33 (3.22, 3.44)\\
\hspace{1em}Regression & $\beta_{1}$ & -3.08 & -3.3 (-3.73, -2.87) & -0.75 & -0.72 (-0.87, -0.52) & 2.6 & 2.5 (2.39, 2.6)\\
\hspace{1em} & $\beta_{2}$ & -2.97 & -3.18 (-3.58, -2.76) & -0.45 & -0.44 (-0.58, -0.26) & 3.43 & 3.42 (3.31, 3.53)\\
\hspace{1em} & $\beta_{3}$ & -2.91 & -3.08 (-3.49, -2.68) & -0.66 & -0.68 (-0.83, -0.48) & 3.04 & 2.98 (2.87, 3.09)\\
\addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{ }}\\
\hspace{1em} & $\sigma_{11}$ & 1 & 0.95 (0.84, 1.02) & 1 & 1 (0.89, 1.11) & 1 & 1.06 (0.97, 1.16)\\
\hspace{1em} & $\sigma_{12}$ & 0.74 & 0.7 (0.59, 0.76) & 0.68 & 0.68 (0.59, 0.78) & -0.45 & -0.41 (-0.47, -0.36)\\
\hspace{1em} & $\sigma_{13}$ & 0.74 & 0.69 (0.58, 0.75) & -0.16 & -0.13 (-0.2, -0.06) & 0.82 & 0.88 (0.79, 0.97)\\
\hspace{1em} & $\sigma_{14}$ & 0.98 & 0.93 (0.81, 0.99) & 0.64 & 0.65 (0.56, 0.75) & 0.7 & 0.75 (0.67, 0.83)\\
\hspace{1em} & $\sigma_{22}$ & 1 & 0.94 (0.82, 1.01) & 1 & 1.03 (0.93, 1.13) & 1 & 1.07 (0.99, 1.16)\\
\hspace{1em} & $\sigma_{23}$ & 0.83 & 0.79 (0.67, 0.85) & -0.43 & -0.4 (-0.46, -0.34) & -0.66 & -0.62 (-0.68, -0.57)\\
\hspace{1em} & $\sigma_{24}$ & 0.81 & 0.77 (0.66, 0.83) & 0.63 & 0.67 (0.58, 0.77) & 0.01 & 0.07 (0.01, 0.13)\\
\hspace{1em} & $\sigma_{33}$ & 1 & 0.96 (0.84, 1.03) & 1 & 1 (0.91, 1.09) & 1 & 1.05 (0.96, 1.15)\\
\hspace{1em} & $\sigma_{34}$ & 0.85 & 0.81 (0.69, 0.87) & 0.15 & 0.15 (0.08, 0.23) & 0.59 & 0.64 (0.56, 0.72)\\
\hspace{1em} & $\sigma_{44}$ & 1 & 0.95 (0.83, 1.01) & 1 & 1.02 (0.92, 1.13) & 1 & 1.06 (0.97, 1.15)\\
\addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{ }}\\
\hspace{1em} & $\psi_{1}$ & -0.33 & -0.33 (-0.62, 0.69) & 0.67 & 0.7 (0.46, 0.89) & -1 & -1.01 (-1.13, -0.87)\\
\hspace{1em} & $\psi_{2}$ & -0.33 & -0.32 (-0.61, 0.64) & 0.67 & 0.63 (0.38, 0.82) & -1 & -0.88 (-1.01, -0.75)\\
\hspace{1em} & $\psi_{3}$ & -0.33 & -0.33 (-0.61, 0.69) & 0.67 & 0.64 (0.43, 0.82) & -1 & -1.01 (-1.14, -0.88)\\
\hspace{1em} & $\psi_{4}$ & -0.33 & -0.31 (-0.63, 0.67) & 0.67 & 0.7 (0.45, 0.89) & -1 & -0.94 (-1.07, -0.81)\\
\addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{ }}\\
\hspace{1em}Multinom. & $\delta_{11}$ & 0.9 & 0.88 (0.81, 0.95) & 0.9 & 0.88 (0.81, 0.95) & 0.9 & 0.88 (0.81, 0.95)\\
\hspace{1em} & $\delta_{12}$ & 0.23 & 0.22 (0.14, 0.3) & 0.23 & 0.22 (0.14, 0.3) & 0.23 & 0.22 (0.14, 0.3)\\
\addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{ }}\\
\hspace{1em}Clustering & $\pi_l$ & 0.28 & 0.28 (0.27, 0.28) & 0.42 & 0.43 (0.42, 0.43) & 0.3 & 0.3 (0.3, 0.3)\\
\bottomrule
\end{tabular}
\end{table}
